Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-22 latex/main.tex
Updates

\section*{INTRODUCTION}
% Classical phylo methods fundamentally do not apply to SARS-CoV-2 data, both
% in terms of the basic goal, assumptions about samples and how recombination
% can be treated.
Classical phylogenetics is largely concerned with inferring the evolutionary history 
of species~\cite{Felsenstein2004}, with 
% Not sure about this assertion! Yan?
samples separated by millions of years of evolution.
[This is rough outline text]
Complex models are required to fill in the gaps in our knowledge,
incorporating prior information about sequence evoltion etc etc.
In this context, recombination can be treated as a nuisance factor 
because we cannot hope to observe it over these timescales [or something
similar].
Phylogenetics has continued along this vein, 
although with various challenges on how to accomodate 
more closely related samples and to handle potential recombination met.

>>>>

\section*{INTRODUCTION}
% Classical phylo methods fundamentally do not apply to SARS-CoV-2 data, both
% in terms of the basic goal, assumptions about samples and how recombination
% can be treated.
Classical phylogenetics is largely concerned with inferring the evolutionary history 
of species~\cite{Felsenstein2004}, 
where recombination
is a problem. 
In this context, recombination is something to be minimised,
with species trees inferred over 
non-recombining regions~\cite{Rannala2025recombination}.

Modern pathogen datasets are not sparsely sampled from different species,
but many isolates from the same species~\cite{Cryptic2022data}
% Cryptic used neighbor-joining tree
% A pairwise genetic distance matrix was constructed for 15,211 isolates by
% comparing pairs of regenotyped VCF files (see “Genomic data processing and
% variant calling” section above for more details). A neighbourhood-joining tree
% was constructed from the distance matrix using quicktree [53].

[This is rough outline text]
Complex models are required to fill in the gaps in our knowledge,
incorporating prior information about sequence evoltion etc etc.
In this context, recombination can be treated as a nuisance factor 
because we cannot hope to observe it over these timescales [or something
similar].
Phylogenetics has continued along this vein, 
although with various challenges on how to accomodate 
more closely related samples and to handle potential recombination met.
